---
title: "EDS 221 Day 4 Computer Session 1"
author: "Anna Ramji"
format: html
editor: visual
---

## Functions

make a function that doubles a value.

```{r}
double_value <- function(my_dollars) {
  print(2 * my_dollars)
}
```

```{r}
double_value(2)
double_value(my_dollars = 10.13)
```

```{r}
# returns error bc only expecting one numeric argument
#double_value(my_dollars = 10.13, 5)
```

```{r}
monthly_income <- c(42.50, 61.75, 89.50)

# uses function on every value in the vector 

double_value(monthly_income)
```

Create a function that returns an awesome restaurant name given a cool animal and your favorite food.

```{r}
name_restaurant <- function(animal, food) {
  return(paste0(animal, food))
}
```

```{r}
name_restaurant(animal = "Hawk", food = "Pizza")
```

Given temperature in degrees Fahrenheit, write a function that converts it to Celsius.

(°F − 32) × 5/9 = 0°C

```{r}
f_to_c <- function(f_temp) {
  converted_temp = round((((f_temp - 32) * 5) / 9), digits = 2)
  return(paste(f_temp, "degrees Fahrenheit is",
                 converted_temp, "degrees Celsius"))
}
```

```{r}
f_to_c(f_temp = 52)
f_to_c(100)
```

OR

```{r}
convert_f_to_c <- function(temp_f) {
  converted_temp = round((((temp_f - 32) * 5) / 9), digits = 2)
  print(converted_temp)
}
```

```{r}
convert_f_to_c(temp_f = 52)
```

Create a function that given inputs of age and hometown, returns "I am \_\_ years old, and I'm from \_\_"

```{r}
about_me <- function(age, hometown) {
  print(paste0("I am ", age, " years old, and I'm from ", hometown, "."))
}
```

```{r}
about_me(23, "San Anselmo")
```

We don't always want to print immediately.

We also don't always need to say "return" in R

-   it's storing the value for you automatically

```{r}
double_value_2 <- function(dollars) {
  dollars * 2
}
```

```{r}
double_value_2(dollars = 10) * 0.08
```

#### Print statements and return statements

Print: "show me", Return: "store this so I can use it later"

## 
